---
title: "Data Cleaning Workshop"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

Practise with data cleaning techniques in this workshop. Let's first install and load `tidyverse` package and import our dataset which was downloaded from `ISLR` package in R. For more information about the data file, you can check this [link](https://rdrr.io/cran/ISLR/man/Credit.html).

To import the dataset, you can either use `read_csv()` (part of the `tidyverse` package) or `read.csv()` (is a built-in R function).


```{r  message=FALSE}

#If tidyverse package is not installed, install it first. 
install.packages("tidyverse")

# Load tidyverse package
library(tidyverse)

# Import our data and save it as crdata
crdata <- read.csv("Credit.csv")

```

Let's do the following tasks:

1. View the first four rows of the dataset.
2. Check the structure of the dataset. Are the data types (or measurement levels) correct?
3. Check the summary of the dataset. In which columns we have missing values?

```{r  message=FALSE}

# View the first four rows with head()
head(crdata, 5)  

# Check the structure of the variables in the dataset by using str() function
str(crdata)

# Check the summary of the dataset
summary(crdata)

```

Before handling the missing values, let's check "Age" column. Do you see anything strange?

Let's check the box plot. We will use `boxplot()` function to perform this task.

```{r message=FALSE}

# Let's have a look at Age column of crdata with boxplot
boxplot(crdata$Age)

```

As it is seen from the box plot, we have outliers in "Age" column of our dataset.
In the next task, we will remove these outliers.

4. Find the indices of records(rows) with `crdata$Age >= 100` and assign these indices to *outliers*. Print the data records with outliers.

```{r message=FALSE}

# Find the indices of records with age >= 100 and assign these indices to outliers
outliers <- which(crdata$Age >= 100)  

# Print data records with outliers 
print(crdata[outliers,])

```

5. Remove data records with `crdata$Age >= 100`. To remove data records in specific rows, you can use the following syntax:

       dataframe[-rowindex,]

This code will remove the rows stored in *rowindex* from the *dataframe*.

```{r message=FALSE}

# Remove data records with age >= 100 
crdata <- crdata[-outliers,] 

```

6. Check the new box plot. 

```{r message=FALSE}

# check the boxplot for Age column 
boxplot(crdata)

```

7. Find the average income by excluding the missing values. Use `mean()` function with argument `na.rm = T`. Save result as *avg_income*.

```{r message=FALSE}

# Find the average income by excluding the missing values
avg_income <- mean(crdata$Income, na.rm = T)

#print average income
print(avg_income)

```

8. Replace missing values in "Income" column with the average income using `replace_na()` function.

```{r message=FALSE}

# Replace missing values in Income column with average income
crdata <- replace_na(crdata, list(Income = avg_income))

```

9. Remove data records with missing values by using `na.omit()` function. Save the result as *crdata_removed*.

```{r message=FALSE}

# Remove records with missing values and assign it to crdata_removed
crdata_removed <- na.omit(crdata)

```

10. Check the summary of the dataset again.

```{r message=FALSE, eval=FALSE}

# Check the summary of the dataset
summary(crdata)

```

11. Calculate the number of records removed from the dataset

```{r message=FALSE, eval=FALSE}

# Calculate the number of removed records
nrow(crdata)-nrow(crdata_removed)

```



